### 2017-07-05 assets/list_of_players.tex
Write list of players.

\item $\phi$ - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item $\pi$ - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item $e$ - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item ALLCorALLD - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{axelrodproject}
\item Adaptive - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Li2011}
\item Adaptive Pavlov 2006 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{kendall2007iterated}
\item Adaptive Pavlov 2011 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Li2011}
\item Adaptive Tit For Tat: 0.5 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Tzafestas2000}
\item Aggravater - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Alternator - \textit{Deterministic} - \textit{Memory depth}: 1. \cite{Axelrod1984, Mittal2009}
\item Alternator Hunter - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Anti Tit For Tat - \textit{Deterministic} - \textit{Memory depth}: 1. \cite{Hilbe2013}
\item AntiCycler - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Appeaser - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Arrogant QLearner - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Average Copier - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Better and Better - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{Prison1998}
\item Bully - \textit{Deterministic} - \textit{Memory depth}: 1. \cite{Nachbar1992}
\item Calculator - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{Prison1998}
\item Cautious QLearner - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item CollectiveStrategy (\textbf{CS}) - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Li2009}
\item Contrite Tit For Tat (\textbf{CTfT}) - \textit{Deterministic} - \textit{Memory depth}: 3. \cite{Axelrod1995}
\item Cooperator - \textit{Deterministic} - \textit{Memory depth}: 0. \cite{Axelrod1984, Mittal2009, Press2012}
\item Cooperator Hunter - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Cycle Hunter - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Cycler CCCCCD - \textit{Deterministic} - \textit{Memory depth}: 5. \cite{axelrodproject}
\item Cycler CCCD - \textit{Deterministic} - \textit{Memory depth}: 3. \cite{axelrodproject}
\item Cycler CCCDCD - \textit{Deterministic} - \textit{Memory depth}: 5. \cite{axelrodproject}
\item Cycler CCD - \textit{Deterministic} - \textit{Memory depth}: 2. \cite{Mittal2009}
\item Cycler DC - \textit{Deterministic} - \textit{Memory depth}: 1. \cite{axelrodproject}
\item Cycler DDC - \textit{Deterministic} - \textit{Memory depth}: 2. \cite{Mittal2009}
\item DBS: 0.75, 3, 4, 3, 5 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Au2006}
\item Davis: 10 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Axelrod1980}
\item Defector - \textit{Deterministic} - \textit{Memory depth}: 0. \cite{Axelrod1984, Mittal2009, Press2012}
\item Defector Hunter - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Desperate - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{Berg2015}
\item DoubleResurrection - \textit{Deterministic} - \textit{Memory depth}: 5. \cite{Eckhart2015}
\item Doubler - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Prison1998}
\item Dynamic Two Tits For Tat - \textit{Stochastic} - \textit{Memory depth}: 2. \cite{axelrodproject}
\item EasyGo - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Li2011, Prison1998}
\item Eatherley - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{Axelrod1980b}
\item Eventual Cycle Hunter - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Evolved ANN - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Evolved ANN 5 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Evolved ANN 5 Noise 05 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Evolved FSM 16 - \textit{Deterministic} - \textit{Memory depth}: 16. \cite{axelrodproject}
\item Evolved FSM 16 Noise 05 - \textit{Deterministic} - \textit{Memory depth}: 16. \cite{axelrodproject}
\item Evolved FSM 4 - \textit{Deterministic} - \textit{Memory depth}: 4. \cite{axelrodproject}
\item Evolved HMM 5 - \textit{Stochastic} - \textit{Memory depth}: 5. \cite{axelrodproject}
\item EvolvedLookerUp1\_1\_1 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item EvolvedLookerUp2\_2\_2 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Feld: 1.0, 0.5, 200 - \textit{Stochastic} - \textit{Memory depth}: 200. \cite{Axelrod1980}
\item Firm But Fair - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{Frean1994}
\item Fool Me Forever - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Fool Me Once - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Forgetful Fool Me Once: 0.05 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Forgetful Grudger - \textit{Deterministic} - \textit{Memory depth}: 10. \cite{axelrodproject}
\item Forgiver - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Forgiving Tit For Tat (\textbf{FTfT}) - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Fortress3 - \textit{Deterministic} - \textit{Memory depth}: 3. \cite{Ashlock2006b}
\item Fortress4 - \textit{Deterministic} - \textit{Memory depth}: 4. \cite{Ashlock2006b}
\item GTFT: 0.1 - \textit{Stochastic} - \textit{Memory depth}: 1.
\item GTFT: 0.3 - \textit{Stochastic} - \textit{Memory depth}: 1.
\item GTFT: 0.33 - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{Gaudesi2016, Nowak1993}
\item GTFT: 0.7 - \textit{Stochastic} - \textit{Memory depth}: 1.
\item GTFT: 0.9 - \textit{Stochastic} - \textit{Memory depth}: 1.
\item General Soft Grudger: n=1,d=4,c=2 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Gradual - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Beaufils1997}
\item Gradual Killer: ('D', 'D', 'D', 'D', 'D', 'C', 'C') - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Prison1998}
\item Grofman - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{Axelrod1980}
\item Grudger - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Axelrod1980, Banks1990, Beaufils1997, Berg2015, Li2011}
\item GrudgerAlternator - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Prison1998}
\item Grumpy: Nice, 10, -10 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Handshake - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Robson1990}
\item Hard Go By Majority - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Mittal2009}
\item Hard Go By Majority: 10 - \textit{Deterministic} - \textit{Memory depth}: 10. \cite{axelrodproject}
\item Hard Go By Majority: 20 - \textit{Deterministic} - \textit{Memory depth}: 20. \cite{axelrodproject}
\item Hard Go By Majority: 40 - \textit{Deterministic} - \textit{Memory depth}: 40. \cite{axelrodproject}
\item Hard Go By Majority: 5 - \textit{Deterministic} - \textit{Memory depth}: 5. \cite{axelrodproject}
\item Hard Prober - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Prison1998}
\item Hard Tit For 2 Tats (\textbf{HTf2T}) - \textit{Deterministic} - \textit{Memory depth}: 3. \cite{Stewart2012}
\item Hard Tit For Tat (\textbf{HTfT}) - \textit{Deterministic} - \textit{Memory depth}: 3. \cite{PD2017}
\item Hesitant QLearner - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Hopeless - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{Berg2015}
\item Inverse - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Inverse Punisher - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Joss: 0.9 - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{Axelrod1980, Stewart2012}
\item Level Punisher - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Eckhart2015}
\item Limited Retaliate 2: 0.08, 15 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Limited Retaliate 3: 0.05, 20 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Limited Retaliate: 0.1, 20 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item MEM2 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Li2014}
\item Math Constant Hunter - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Meta Hunter Aggressive: 7 players - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Meta Hunter: 6 players - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Meta Mixer: 173 players - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Naive Prober: 0.1 - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{Li2011}
\item Negation - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{PD2017}
\item Nice Average Copier - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Nydegger - \textit{Deterministic} - \textit{Memory depth}: 3. \cite{Axelrod1980}
\item Omega TFT: 3, 8 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{kendall2007iterated}
\item Once Bitten - \textit{Deterministic} - \textit{Memory depth}: 12. \cite{axelrodproject}
\item Opposite Grudger - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item PSO Gambler 1\_1\_1 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item PSO Gambler 2\_2\_2 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item PSO Gambler 2\_2\_2 Noise 05 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item PSO Gambler Mem1 - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{axelrodproject}
\item Predator - \textit{Deterministic} - \textit{Memory depth}: 9. \cite{Ashlock2006b}
\item Prober - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Li2011}
\item Prober 2 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Prison1998}
\item Prober 3 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Prison1998}
\item Prober 4 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Prison1998}
\item Pun1 - \textit{Deterministic} - \textit{Memory depth}: 2. \cite{Ashlock2006}
\item Punisher - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Raider - \textit{Deterministic} - \textit{Memory depth}: 3. \cite{Ashlock2014}
\item Random Hunter - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Random: 0.1 - \textit{Stochastic} - \textit{Memory depth}: 0.
\item Random: 0.3 - \textit{Stochastic} - \textit{Memory depth}: 0.
\item Random: 0.5 - \textit{Stochastic} - \textit{Memory depth}: 0. \cite{Axelrod1980, Tzafestas2000}
\item Random: 0.7 - \textit{Stochastic} - \textit{Memory depth}: 0.
\item Random: 0.9 - \textit{Stochastic} - \textit{Memory depth}: 0.
\item Remorseful Prober: 0.1 - \textit{Stochastic} - \textit{Memory depth}: 2. \cite{Li2011}
\item Resurrection - \textit{Deterministic} - \textit{Memory depth}: 5. \cite{Eckhart2015}
\item Retaliate 2: 0.08 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Retaliate 3: 0.05 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Retaliate: 0.1 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Revised Downing: True - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Axelrod1980}
\item Ripoff - \textit{Deterministic} - \textit{Memory depth}: 2. \cite{Ashlock2008}
\item Risky QLearner - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item SelfSteem - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{Andre2013}
\item ShortMem - \textit{Deterministic} - \textit{Memory depth}: 10. \cite{Andre2013}
\item Shubik - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Axelrod1980}
\item Slow Tit For Two Tats - \textit{Deterministic} - \textit{Memory depth}: 2. \cite{axelrodproject}
\item Slow Tit For Two Tats 2 - \textit{Deterministic} - \textit{Memory depth}: 2. \cite{Prison1998}
\item Sneaky Tit For Tat - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Soft Go By Majority - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Axelrod1984, Mittal2009}
\item Soft Go By Majority: 10 - \textit{Deterministic} - \textit{Memory depth}: 10. \cite{axelrodproject}
\item Soft Go By Majority: 20 - \textit{Deterministic} - \textit{Memory depth}: 20. \cite{axelrodproject}
\item Soft Go By Majority: 40 - \textit{Deterministic} - \textit{Memory depth}: 40. \cite{axelrodproject}
\item Soft Go By Majority: 5 - \textit{Deterministic} - \textit{Memory depth}: 5. \cite{axelrodproject}
\item Soft Grudger - \textit{Deterministic} - \textit{Memory depth}: 6. \cite{Li2011}
\item Soft Joss: 0.9 - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{Prison1998}
\item SolutionB1 - \textit{Deterministic} - \textit{Memory depth}: 3. \cite{Ashlock2015}
\item SolutionB5 - \textit{Deterministic} - \textit{Memory depth}: 5. \cite{Ashlock2015}
\item Spiteful Tit For Tat - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Prison1998}
\item Stochastic Cooperator - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{Adami2013}
\item Stochastic WSLS: 0.05 - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{axelrodproject}
\item Suspicious Tit For Tat - \textit{Deterministic} - \textit{Memory depth}: 1. \cite{Beaufils1997, Hilbe2013}
\item TF1 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item TF2 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item TF3 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Tester - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Axelrod1980b}
\item ThueMorse - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item ThueMorseInverse - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Thumper - \textit{Deterministic} - \textit{Memory depth}: 2. \cite{Ashlock2008}
\item Tit For 2 Tats (\textbf{Tf2T}) - \textit{Deterministic} - \textit{Memory depth}: 2. \cite{Axelrod1984}
\item Tit For Tat (\textbf{TfT}) - \textit{Deterministic} - \textit{Memory depth}: 1. \cite{Axelrod1980}
\item Tricky Cooperator - \textit{Deterministic} - \textit{Memory depth}: 10. \cite{axelrodproject}
\item Tricky Defector - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{axelrodproject}
\item Tullock: 11 - \textit{Stochastic} - \textit{Memory depth}: 11. \cite{Axelrod1980}
\item Two Tits For Tat (\textbf{2TfT}) - \textit{Deterministic} - \textit{Memory depth}: 2. \cite{Axelrod1984}
\item VeryBad - \textit{Deterministic} - \textit{Memory depth}: \(\infty\). \cite{Andre2013}
\item Willing - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{Berg2015}
\item Win-Shift Lose-Stay: D (\textbf{WShLSt}) - \textit{Deterministic} - \textit{Memory depth}: 1. \cite{Li2011}
\item Win-Stay Lose-Shift: C (\textbf{WSLS}) - \textit{Deterministic} - \textit{Memory depth}: 1. \cite{Kraines1989, Nowak1993, Stewart2012}
\item Winner12 - \textit{Deterministic} - \textit{Memory depth}: 2. \cite{Mathieu2015}
\item Winner21 - \textit{Deterministic} - \textit{Memory depth}: 2. \cite{Mathieu2015}
\item Worse and Worse - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{Prison1998}
\item Worse and Worse 2 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{Prison1998}
\item Worse and Worse 3 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\). \cite{Prison1998}
\item ZD-Extort-2 v2: 0.125, 0.5, 1 - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{Kuhn2017}
\item ZD-Extort-2: 0.1111111111111111, 0.5 - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{Stewart2012}
\item ZD-Extort-4: 0.23529411764705882, 0.25, 1 - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{axelrodproject}
\item ZD-GEN-2: 0.125, 0.5, 3 - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{Kuhn2017}
\item ZD-GTFT-2: 0.25, 0.5 - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{Stewart2012}
\item ZD-SET-2: 0.25, 0.0, 2 - \textit{Stochastic} - \textit{Memory depth}: 1. \cite{Kuhn2017}